### 2019-03-07 manuscript/introduction.tex
various minor manuscript edits

%B: the key alteration below is an attempt at easing the transitions of the complicated existing logical structure, which intermixes advances with setbacks and problems.
%B: the "Here, we" part should maybe come at the bottom of introduction, and it's a little early for here, in the first paragraph.
Species accumulate around the tree of life at different rates.
A possible explanation for this phenomenon is that species possess various traits or character states that may differentially affect rates of diversification. %B: new.
Dramatic increases in available phylogenetic and phenotypic data %citation
as well as methodological advances %citation
have greatly accelerated the search for such traits that influence diversification traits. Nevertheless, identifying focal traits associated with rates of speciation and extinction remains as a difficult statistical problem (see \citep{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}) . %B: the following is now orphaned, because it's inclusion merits addition of non-trait-focused methods; "potentially interesting biological processes"
In part, this is a difficult problem because speciation and extinction rarely depend on a single trait; the context  in which traits occurs can lead to complex interactions resulting in a heterogenous speciation and extinction rates (\citep{beaulieu_2016, caetano_2018, herrera_2018}). Consequently, examining the association of only one trait when multiple complex interactions are changing diversification patterns can be misleading.

@@

Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
In particular, self-incompatibility systems cause a plant to reject its own pollen, and their loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution. % refs: stebbins, igic
Previous phylogenetic analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014}, but they have not considered the possibility of other correlated traits driving this pattern.
Given that changes in ploidy and breeding systems may be causally related and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
This includes their joint influence on lineage diversification, and also potential patterns in the order of their transitions.
For example, do losses of SI more commonly occur tied to polyploidization, or without a ploidy shift?
Do polyploids arise more commonly from self-incompatible or self-compatible diploids?
\Citet{robertson_2011} found that the pathway from self-incompatible diploids to self-compatible polyploids is dominated by loss of self-incompatibility followed later by polyploidization over long timescales, but proceeds in one step via polyploidization of SI species over short timescales.
Greatly improved phylogeny and methods that allow for diversification rate differences allow far more powerful approach to such problems. % E: Technically, we allow for div effects in the rate estimates, but not in the pathways.  If SC-D have so much extinction that the step to SC-P never happens, the old pathway methods won't see that.  I'll think about if this is easy to fix.
% And, are diversification rates of polyploid SC lineages different if one or other path was taken? \citet{charlesworth1985} % E: This is a great question.  We don't answer it with this round of analyses.  Maybe when we have more time for the revision?
% E: Relatedly, I'm wondering if pathways should be a separate paper.  There is the robertson2011 question, the modification that allows for diversification, and the charlesworth1985 question.  And dioecy.
%B: Yes, agreed.

>>>>

%B: the key alteration below is an attempt at easing the transitions of the complicated existing logical structure, which intermixes advances with setbacks and problems.
%B: the "Here, we" part should maybe come at the bottom of introduction, and it's a little early for here, in the first paragraph.
Species accumulate around the tree of life at different rates.
A possible explanation for this phenomenon is that species possess various traits or character states that may differentially affect rates of diversification. %B: new.
Dramatic increases in available phylogenetic and phenotypic data %citation
as well as methodological advances %citation
have greatly accelerated the search for such traits that influence diversification traits. Nevertheless, identifying focal traits associated with rates of speciation and extinction remains a difficult statistical problem (see \citep{maddison_2015, rabosky_2015, moore_2016, fitzjohn_2009, goldberg_2012, beaulieu_2016, rabosky_2017}) . %B: the following is now orphaned, because it's inclusion merits addition of non-trait-focused methods; "potentially interesting biological processes"
In part, this is a difficult problem because speciation and extinction rarely depend on a single trait; the context  in which traits occurs can lead to complex interactions resulting in a heterogenous speciation and extinction rates (\citep{beaulieu_2016, caetano_2018, herrera_2018}). Consequently, examining the association of only one trait when multiple complex interactions are changing diversification patterns can be misleading.

@@

Breeding system shifts---changes in the collection of physiological and morphological traits that determine the likelihood that any two gametes unite---are remarkably common and affect the distribution and amount of genetic variation in populations \citep{stebbins1974, barrett2013}.
In particular, self-incompatibility systems cause a plant to reject its own pollen, and their loss, yielding self-compatibility, is one of the most replicated transitions in flowering plant evolution. % refs: stebbins, igic
Previous phylogenetic analyses have reported higher rates of diversification for self-incompatible than for self-compatible lineages \citep{goldberg_2010, devos2014, freyman_2018}, but they did not test other correlated traits driving this pattern.
Given that changes in ploidy and breeding systems may be causally related and have profound affects on the fate of lineages, it seems particularly profitable to examine possible interactions in their macroevolutionary effects.
This includes their joint influence on lineage diversification, and also potential patterns in the order of their transitions.
For example, do losses of SI more commonly occur tied to polyploidization, or without a ploidy shift?
Do polyploids arise more commonly from self-incompatible or self-compatible diploids?
\Citet{robertson_2011} found that the pathway from self-incompatible diploids to self-compatible polyploids is dominated by loss of self-incompatibility followed later by polyploidization over long timescales, but proceeds in one step via polyploidization of SI species over short timescales.
Greatly improved phylogeny and methods that allow for diversification rate differences allow far more powerful approach to such problems. % E: Technically, we allow for div effects in the rate estimates, but not in the pathways.  If SC-D have so much extinction that the step to SC-P never happens, the old pathway methods won't see that.  I'll think about if this is easy to fix.
% And, are diversification rates of polyploid SC lineages different if one or other path was taken? \citet{charlesworth1985} % E: This is a great question.  We don't answer it with this round of analyses.  Maybe when we have more time for the revision?
% E: Relatedly, I'm wondering if pathways should be a separate paper.  There is the robertson2011 question, the modification that allows for diversification, and the charlesworth1985 question.  And dioecy.
%B: Yes, agreed.

### 2019-03-07 manuscript/methods.tex
pathway note
various minor manuscript edits

To assess the effects of breeding system in the diversification process, we first fit model in which the states are self-incompatible (I) or self-compatible (C).
This is the same as the analysis of \citet{goldberg_2010}, save for an updated phylogeny \citep{sarkinen_2013}.
We call this BiSSE model the \textit{I/C breeding system} model. 
To parse out the effect of breeding system on diversification, while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
We call this HiSSE model the \textit{I/C+A/B breeding system and hidden state model}.

@@

For all four models that consider ploidy changes, we allowed diploidization by adding a parameter $\delta$ connecting polyploidy state back to diploid or to self-compatible diploid in the mult-state models (denoted by $+/delta$ in the models of  \cref{table:marginallike}).
Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models like chromEvol \citep{mayrose_2010}.
These types of classifications do not allow for a ploidy reversion.
Where indicated, the classification of ploidy for the data used in our models was based on chromosome multiplicity at the genus level.
% ref data/sources table?
However, the majority of the ploidy classifications were adopted from original studies with alternative sources of information (\eg geographic distribution, genus ploidy distribution) where ploidy was defined by authors that found evidence for it.
Since it is not clear whether diploidization can be detected under alternative ploidy classifications or even classifications based on chromosome number multiplicity at the genus level, we also fit the models without diploidization in order to test  whether the conclusions about diversification are sensitive to including diploidization.
As discussed by \citet{servedio_2014}, the presence or absence of a hypothesis can have an exploratory goal.
In our case the diploidization parameter (or its absence, $\delta=0$) in our models is an opportunity to explore an assumption that might be important but that is not the single definitive process to understand the interactions among polyploidy, breeding system, and diversification.

@@

Parameters for each of the 10 diversification models were estimated using custom code in the RevBayes \citep{hoehna_2016} environment.
Code for analyses and key results is available at \url{https://github.com/roszenil/solploidy}.
We included a correction for incomplete sampling in all analyses, based on assuming that the Solanaceae family has approximately 3,000 species ($s=595/3000$) as estimated by the Solanaceae Source project \citep{solsource}.
For all 10 models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
For each model, an MCMC chain was run for 96 hours in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the 10 models. %B: the number of models changes here? R-my mistake! fixed it
For each model, convergence and mixing of the MCMC was tested using the R library \texttt{coda} and the software package Tracer (see supplementary information for convergence plots). % fixme: eventually, add supp info fig ref

% R will add a bit about the ASR methods
In the RevBayes code we calculated the probability of ancestral states at each node as performed chromosome number change models proposed in  \citet{freyman_2017}. Ancestral reconstructions for all models can be found in the supplementary information and show the maximum a posteriori of the marginal probability distributions for each of the 594 internal nodes for each of the 10 models.

>>>>

To assess the effects of breeding system in the diversification process, we first fit model in which the states are self-incompatible (I) or self-compatible (C).
This is the same as the analysis of \citet{goldberg_2010}, save for an updated phylogeny \citep{sarkinen_2013}.
We call this BiSSE model the \textit{I/C breeding system} model. 
To parse out the effect of breeding system on diversification, while allowing for the possibility of heterogeneous diversification rates unrelated to breeding system, we subdivided each of those states into hidden states $A$ and $B$.
We call this HiSSE model the \textit{I/C+A/B breeding system and hidden state model}.
This model was previously used in \citet{freyman_2018} to examine the effect of
breeding system on diversification in the plant family Onagraceae.

@@

For all four models that consider ploidy changes, we allowed diploidization by adding a parameter $\delta$ connecting polyploidy state back to diploid or to self-compatible diploid in the mult-state models (denoted by $+/delta$ in the models of  \cref{table:marginallike}).
Previous modeling approaches \citep{mayrose_2011} have argued against inferring diploidization rates when using ploidy data that comes from classifications based on chromosome number multiplicity or chromosome number change models \citep{mayrose_2010, glick2014chromevol, freyman_2017}.
These types of classifications do not allow for a ploidy reversion.
Where indicated, the classification of ploidy for the data used in our models was based on chromosome multiplicity at the genus level.
% ref data/sources table?
However, the majority of the ploidy classifications were adopted from original studies with alternative sources of information (\eg geographic distribution, genus ploidy distribution) where ploidy was defined by authors that found evidence for it.
Since it is not clear whether diploidization can be detected under alternative ploidy classifications or even classifications based on chromosome number multiplicity at the genus level, we also fit the models without diploidization in order to test  whether the conclusions about diversification are sensitive to including diploidization.
As discussed by \citet{servedio_2014}, the presence or absence of a hypothesis can have an exploratory goal.
In our case the diploidization parameter (or its absence, $\delta=0$) in our models is an opportunity to explore an assumption that might be important but that is not the single definitive process to understand the interactions among polyploidy, breeding system, and diversification.

@@

Parameters for each of the 10 diversification models were estimated using RevBayes \citep{hoehna_2016}.
Code for analyses and key results are available at \url{https://github.com/roszenil/solploidy}.
We accounted for incomplete sampling in all analyses by fixing the probability of sampling a species at the present $s$ to $595/3000$ since the Solanaceae family has approximately 3,000 species as estimated by the Solanaceae Source project \citep{solsource}.
For all 10 models, we assumed that speciation and extinction parameters had log-normal prior distributions with means equal to the expected net diversification rate $(\text{number of taxa} / [2 \times \text{root age}])$ and standard deviation $0.5$.
Priors for parameters defining trait changes were assumed to be gamma distributed with parameters $k=0.5$ and $\theta=1$. 
For each model, a Markov chain Monte Carlo \citep[MCMC;][]{metropolis1953equation,Hastings1970} was run for 96 hours in the high-performance computational cluster at the Minnesota Supercomputing Institute, which allowed for 5,000 generations of burn-in and a minimum of 200,000 generations of MCMC for each of the 10 models. %B: the number of models changes here? R-my mistake! fixed it
Convergence and mixing of each MCMC 
was determined by ensuring the effective sample size of each parameter was over 200.

For each model, ancestral states at each node in the phylogeny were sampled jointly during the MCMC analyses. Ancestral state reconstructions for all models can be found in the supplementary information and show the maximum a posteriori of the marginal probability distributions for each of the 594 internal nodes for each of the 10 models.
